---
title: "exam3"
format:
  html:
    embed-resources: true
editor: visual
---

---

---
title: "exam3"
format:
  html:
    embed-resources: true
---

## 1

After using the setwd() command to set the wd to the Exam_3 folder:

```{r}
sal<-read.csv("FacultySalaries_1995.csv")
```

converting into long format

```{r}
library(tidyverse)

sal_tidy <- sal %>%
  pivot_longer(
    cols = c(AvgFullProfSalary, AvgAssocProfSalary, AvgAssistProfSalary),
    names_to = "Rank",
    values_to = "Salary"
  ) %>%
  mutate(
    Rank = recode(Rank,
                  "AvgFullProfSalary" = "Full",
                  "AvgAssocProfSalary" = "Associate",
                  "AvgAssistProfSalary" = "Assistant")
  )

```

\
Filter VIIB out from dataset

```{r}
sal_tidy_filtered <- sal_tidy %>%
  filter(Tier != "VIIB")
```

create plot where Tier is sorted by color\

```{r}
ggplot(sal_tidy_filtered, aes(x = Rank, y = Salary, fill = Tier)) +
  geom_boxplot() +
  labs(title = "Salary by Rank (Colored by Tier)")
```

# 2

```{r}
model <- aov(Salary ~ State + Tier + Rank, data = sal_tidy_filtered)

summary(model)
```

# 3

```{r}
  juniper<-read.csv("Juniper_Oils.csv")
```

Y is concentration, X is years since burn, and we want the graph to be faceted according to each Chemical ID using free Y-axis scales

Convert to long format

```{r}
juniper_long <- juniper %>%
  pivot_longer(
    cols = -c(SampleID, Project, Amplicon, Tree_Species, BurnYear,
              Latitude, Longitude, Field_Office, BLM_Fire_Name,
              Tracking_Number, Yield_percent, Bolt_Surface_Area_cm2,
              Raw_Exit_Holes_per_cm2, Raw_Exit_Holes,
              Living_Larvae, ChemTotal, ChemMean, YearsSinceBurn),
    names_to = "Compound",
    values_to = "Concentration"
  )
```

# 4

graphing it using ggplot

```{r}
library(ggplot2)
ggplot(juniper_long, aes(x = YearsSinceBurn, y = Concentration)) +
  geom_line() +
  geom_smooth(span = 0.4) +
  facet_wrap(~Compound,scales = "free_y")
```

# 5

Load pkg's

```{r}

library(tidyverse)
library(broom)
```

filter out nonpositives for gamma model requirement\
\

```{r}
juniper_long <- juniper_long %>%
  filter(Concentration > 0)
```

Create a generalized linear model where concentration is analyzed as a function of

```{r}
results <- juniper_long %>%
  group_by(Compound) %>%
  do(broom::tidy(glm(Concentration ~ YearsSinceBurn,
                     data = .,
                     family = Gamma(link = "log"))))

results <- results %>%
  filter(term == "YearsSinceBurn")
```

\
filter data frame to isolate statistically significant p-values

```{r}
sig_results <- results %>%
  filter(p.value < 0.05)
```
